---
title: "Report"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Visualize and fill missing values
library("Amelia")
# Visualize outliers
library("mvoutlier")
# Easier manipulation
library("dplyr")
# Plot tools
library("ggplot2")
# Correlation matrix
library("Hmisc")
# Variety of models
library("caret")
library("e1071")
# Plot decision tree
library("rattle")
library("rpart.plot")
library("RColorBrewer")

data<-read.csv("/Users/thomas-legrand/Documents/DCU/Data Analytics/Assignment-1/iris_with_missing_data.csv", sep=",")
```

## Problem recognition

We want to be able to know what species a flower belongs to, according to 4 parameters concerning the sepal and the petal.

## Outlook of the data

```{r head, echo=FALSE}
head(data)
data(data)
```

We notice that some data are missing.
```{r missmap, echo=FALSE}
missmap(data)
```

Thanks to the Amelia package, we can fill the missing data using the expectation–maximization (EM) algorithm.
```{r amelia, include=FALSE}
a.out<-amelia(data, m = 1, idvars='Species')
amelia<-write.amelia(obj=a.out, file.stem = "amelia")
amelia1<-read.csv("~/Documents/DCU/Data Analytics/Assignment-1/amelia1.csv")
amelia1<-select(amelia1, Sepal.Length:Species)
```


Let's split randomly the data into 2 parts : a training and a testing dataset.
By default, createDataPartition does a stratified random split of the data.
```{r separate}
inTrain = createDataPartition(amelia1$Species, p = 2/3, list = FALSE)
dfTrain=amelia1[inTrain,]
dfTest=amelia1[-inTrain,]
```

## Exploration of the data 

```{r describe, echo=FALSE}
summary(dfTrain)
plot(dfTrain)
```

## Analysis

### Comparison using qqplot

```{r comparison, echo=FALSE}
qplot(Species, Sepal.Length, data=dfTrain, geom=c("boxplot", "jitter"), 
   fill=Species)
qplot(Species, Sepal.Width, data=dfTrain, geom=c("boxplot", "jitter"), 
   fill=Species)
qplot(Species, Petal.Length, data=dfTrain, geom=c("boxplot", "jitter"), 
   fill=Species)
qplot(Species, Petal.Width, data=dfTrain, geom=c("boxplot", "jitter"), 
   fill=Species)
```

We can notice that the petal characteristics are more discriminant than the sepal ones.
Indeed, the setosa is way different than the 2 others species.

### Looking for correlation
```{r correlation, echo=FALSE}
# Hmisc package
rcorr(as.matrix(select(dfTrain, Sepal.Length:Petal.Width)))
```

It is clear that the petal width and the petal length are higly corrolated. 
The sepal length is corrolated with the petal length and width.

### Classification
We want to classify the data in 3 classes. 

#### Decision Tree

```{r tree, echo=FALSE}
tree.fit <- rpart(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, method="class", data=dfTrain)
tree.model <- predict(tree.fit, dfTest, type="class")
fancyRpartPlot(tree.fit)
```

#### Naive Bayes Classifier

```{r naive}
bayes.fit<-naiveBayes(dfTrain[,1:4], dfTrain[,5]) 
bayes.model<-predict(bayes.fit, dfTest[,-5])
```

#### SVM

```{r tune, echo=FALSE}
tune.svm(Species~., data = dfTrain, gamma = 10^(-6:-1), cost = 10^(-1:1))
```

Let's build a svm model with a radial kernel to predict species using C=10 and gamma=0.01, which were the best values according the tune() function run before.

```{r svm}
svm.model<-svm(Species~., data=dfTrain, gamma = 0.01, cost = 10) 
svm.pred <- predict(svm.model, dfTest[,-5])
```


## Results

### Decision Tree

```{r resultsTree, echo=FALSE}
(table(tree.model, dfTest$Species)/16)*100
```

We can see that despite its simplicity the results are quite good.
As we expected, the difficulty was about distinguishing versicolor and virginica.

### Naive Bayes Classifier

```{r resultsNB, echo=FALSE}
(table(bayes.model, dfTest[,5])/16)*100
```

The Naive Bayes classifier seems to have better performance.

### SVM

```{r resultsSvm, echo=FALSE}
(table(pred = svm.pred, true = dfTest[,5])/16)*100
```